---
title: "Air quality project : Time Series Analysis"
author : Yousra Cherkaoui Tangi - Lou Peltier - Guilhem Bois
output: pdf_document
date : "`r format(Sys.time(), '%d %B, %Y')`"
---

\tableofcontents
\newpage
\section{Preliminary }

Our project consists of time series analysis using R programming language. Our data set contains the responses of a gas mutisensor device deployed on the field in an Italian city, in a significantly polluted area, at road level. The data set contains 9358 instances of hourly averaged responses from an array of 5 metal oxide chemical sensors embedded in an Air Quality Chemical Multisensor Device. Data were recorded from March 2004 to February 2005. 

```{r, echo=FALSE}
#Spécifier le chemin du fichier .csv 
fic = "/home/yousra/3A/cours/SeriesTemp/Air-Quality-Project/data/AirQualityUCI.csv"
data = read.csv(fic, header = TRUE, sep = ";", stringsAsFactors = FALSE)
```


```{r}
dim(data)
str(data)
```
| Attribute 	    	|Description                                                                                           |
|:-----------------:|:---------------------------------------------------------------------------------------------------- |    
| Date 		          | Date (DD/MM/YYYY)                                                                                    |
| Time 		          | Time (HH.MM.SS)                                                                                      |
| CO(GT) 		        | True hourly averaged concentration CO in mg/m^3 (reference analyzer)                                 |
| PT08.S1(CO)     	| PT08.S1 (tin oxide) hourly averaged sensor response (nominally CO targeted)                          |
| NMHC(GT) 		      | True hourly averaged overall Non Metanic Hydro Carbons concentration in microg/m^3(reference analyzer |
| C6H6(GT) 		      | True hourly averaged Benzene concentration in microg/m^3 (reference analyzer)                        |
| PT08.S2(NMHC)     | PT08.S2 (titania) hourly averaged sensor response (nominally NMHC targeted)                          |
| NOx(GT) 		      | True hourly averaged NOx concentration in ppb (reference analyzer)                                   |
| PT08.S3(NOx)    	| PT08.S3 (tungsten oxide) hourly averaged sensor response (nominally NOx targeted)                    |
| NO2(GT) 		      | True hourly averaged NO2 concentration in microg/m^3 (reference analyzer)                            |
| PT08.S4(NO2)    	| PT08.S4 (tungsten oxide) hourly averaged sensor response (nominally NO2 targeted)                    |
| PT08.S5(O3)     	| PT08.S5 (indium oxide) hourly averaged sensor response (nominally O3 targeted)                       |
| T 		            | Temperature in °C                                                                                   |
| RH 		            | Relative Humidity (%)                                                                                |
| AH 		            | AH Absolute Humidity                                                                                 |
 		

\section{I)-Cleaning and preprocessing}
As we can see in the str(dim) output, our data contains a certain number of variables that are of character type which make it not directly exploitable, also some numbers

\subsection{1) Data cleaning}s

\subsection{2) Data preprocessing}

\section{II)-Model fitting on the time series of interest}

\subsection{1) MA model}

\subsection{2) AR model}


\subsection{3) ARMA model}

\subsection{4) Residuals} 

\subsection{5) GARCH model}

\subsection{6) Prediction intervals for the 10 most recent data}

\section{III)-Training on the times series of interest using explanatory times series}

\subsection{1) Preprocessing}

\subsection{2) Time varying coefficients}

\subsection{3) QLIK}

\subsection{4) Prediction}

\section{IV) Conclusion}